\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-10-01 papers/methods/paper.tex
Updating paper

% Abstract of the paper
\begin{abstract}
We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects. This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improving treatment of Malmquist bias, whilst increasing numerical efficiency. Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery. We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample. SNANA simulations can recover the input cosmology {\green between $0$ and $0.5\sigma$ bias depending on the intrinsic scatter model} for the DES 3-year spectroscopic sample, while taking into account an increased number of systematics and increasing simulation independent over previous analyses. {\green Future development will focus on increasing model agnosticism towards intrinsic scatter model.} {\red Have left out explicit references for `previous analyses' as I dont want to bloat the abstract. Its talked about in the paper, so is this alright?}
\end{abstract}
% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. {\green This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}.} {\red  Ive just swapped out references here as this isnt a critical point for my analysis to drill down on.} The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable. {\red Removed confusing statement on intrinsic dispersion.}

@@

Whilst there are a large number of hierarchical models available, all so far mentioned, in the case where any validation has been done, have been done on either $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology. None of them have undergone high-statistics simulation verification for $w$CDM cosmology to quantify each models' respective bias, which is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$. {\red Dan, you mention underselling the prize here. Any suggestions on an appropriate sell, so to speak?}

@@

Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. As the SALT2 model does not include {\green full treatment of} intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$. These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

@@

\begin{table}
	\centering
	\caption{ Parameters defined in our model and a summary of their use. The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova. The bottom section shows the observables (not parameters) that are the input data to the model.}
	\label{tab:param_summary}
	\begin{tabular}{ll}
		\hline
		Parameter & Description \\
		\hline
		$\Omega_m$  & Matter density  \\
		$w$  &  Dark energy equation of state  \\
		$\alpha$ & Stretch standardisation   \\
		$\beta$         &  Colour standardisation   \\
		$\delta(0)$ & Scale of the mass-magnitude correction\\
		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
		$\delta\mathcal{Z}_i$ & Systematics scale\\
		$\langle M_B \rangle$ & Mean absolute magnitude \\
		%$s_c$ & Colour bias\\
		\hline
		$\delta S$ & Selection effect deviation \\
		$\langle x_1^i \rangle$ & Mean stretch nodes\\
		$\langle c^i \rangle$ & Mean colour nodes\\
		$\alpha_c$ & Skewness of colour population \\
		$\sigma_{M_B}$ & Population magnitude scatter \\
		$\sigma_{x_1}$ & Population stretch scatter \\
		$\sigma_{c}$ & Population colour scatter \\
		$\kappa_{0}$ & Extra colour dispersion\\
		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
		\hline
		$m_B$ & True flux\\
		$x_1$ & True stretch \\
		$c$ & True colour \\
		$z$ & True redshift \\
		$M_B$ & Derived absolute magnitude \\
		$\mu$ & Derived distance modulus \\
		\hline
		$\hat{m}_B$ & Measured flux\\
		$\hat{x}_1$ & Measured stretch \\
		$\hat{c}$ & Measured colour \\
		$C$ & Covariance on flux, stretch and colour \\
		$\hat{z}$ & Observed redshift \\
		$\hat{m}$ & Determined mass probability \\
		$\hat{p}$ & Determined {\green Ia} probability \\
		\hline
	\end{tabular}
\end{table}

@@

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the {\green width of the underlying magnitude population (discussed in Section \ref{sec:underlying})} without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion. {\green Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.}

As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature). The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$. {\green We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$. We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}.} As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.

Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, {\green by modelling the dispersion as a multivariate Gaussian}, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient. We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations. This extra dispersion is modelled with redshift independence. {\red I'm not too sure what you mean about telling a story here. I've tried to change the wording to put it back in line with the method, is that sufficient? Also, I did not try redshift dependent extra scatter on $m_B$.}

@@

We tested two models with varying $\beta$ against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$. See Section \ref{sec:simdes} for details on simulation generation. We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects. This degeneracy resulted in a significant bias in recovered cosmology. {\red Removed comments about varying alpha and beta.} {\green Due to the }recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses. As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.

@@

Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population {\green \citep{Conley2011, Betoule2014}}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.

@@

In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based. {\green We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.}

@@

Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}. The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  {\green and secondary parameters are shown in Table \ref{tab:standardisation}. Our recovered results for $w$ are all within $0.5\sigma$ of the true values of $w=-1$. Table \ref{tab:standardisation} shows a clear difference in both $\beta$ and $\sigma_{m_B}$ across the {\gten} and {\celeven} simulations. As expected, the {\celeven} simulations recover a far smaller intrinsic magnitude scatter, giving results of approximately $0.025$ when compared to the result of $0.070$ for the {\gten} simulations. The extra smearing of the {\celeven} model does not result in a biased $\beta$ value, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$. Interestingly, $w$-bias is only found for the {\celeven} simulations.}

We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics. {\green To confirm this, we run two tests. The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model.} From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.

Unfortunately, adding extra fit parameters to allow for shifting observables {\green washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model), which presents difficulties when trying to account for correlations with population and scatter model.} This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically. Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.

@@

Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics. Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters. {\green As expected, $w$ bias would be introduced primarily through incorrectly determined standardisation parameters, mean absolute magnitude or incorrect dispersion parameters. As discussed previously, the culprit for our $w$-bias is the dispersion model, rather than standardisation parameters}. Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}. The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.

@@

\begin{table*}
	\centering
	\caption{The error budget on $w$, as determined from analysing uncertainty on simulation data whilst progressively enabling model features. We start from the top of the table, only varying cosmological parameters $\Omega_m$ and $w$, and then progressively unlock parameters and let them fit as we progress down the table. The cumulative uncertainty shows the total uncertainty on $w$ on the fit for all, where the $\sigma_w$ term is derived by taking the quadrature difference in cumulative uncertainty as we progress.}
	\label{tab:uncert}
	\begin{tabular}{l|l|rr}
		Feature & {\green Parameters} & $\sigma_w$ & Cumulative  \\
		\hline
		Cosmology                      & $\Omega_m$, $w$ & 0.051  &   0.051       \\
		Standardisation                & $\alpha$, $\beta$, $\langle M_B \rangle$, $\delta(0)$, $\delta(\infty)/\delta(0)$ & 0.046  &   0.068       \\
		Intrinsic scatter              & $\kappa_0$, $\kappa_1$ & 0.020  &   0.071       \\
		Redshift-independent populations             & $\sigma_{M_B}$, $\sigma_c$, $\sigma_{x_1}$, $\alpha_c$ & 0.022  &   0.074       \\
		Redshift-dependent populations & $\langle c_i \rangle$, $\langle x_{1,i} \rangle$ & 0.030  &   0.080   \\
		Systematics                    & $\delta \mathcal{Z}_i$, $\delta S$ & 0.054  &   0.096      \\ 
		\hline
	\end{tabular}
\end{table*}

@@

{\green As shown in \citet{Brout18SYS}, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known. As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}. As such, we expect the BBC method to be biased in one direction for {\gten} simulations and the other direction for {\celeven} simulations. These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}, as both} BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model. The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$. Our method shows a small improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}. This decrease in error is not statistically significant as we have statistical uncertainty of $\sim0.01$ for $100$ simulation realisations. The average bias over the two scatter models is $+0.028$, representing a larger bias than the BBC method.

@@

\begin{table}
	\centering
	\caption{Characterising the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model. The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation. Finally, we also characterise the scatter between the methods in the final row. Averages are computed giving each simulation sample the same weight.}
	\label{tab:bulk_summary_bbc}
	\begin{tabular}{l|ccc}
		\hline
		Model & {\gten} & {\celeven} & Mean ({\gten} + {\celeven}) \\
		\hline
		Steve $\langle w \rangle$           & $-0.998$ & $-0.945$ & $-0.972$\\
		BBC $\langle w \rangle$             & $-1.044$ & $-0.978$ & $-1.010$ \\
		(Steve - BBC)  $\langle w \rangle$  & $+0.044$ & $+0.033$ & $+0.039$ \\
		\hline
		(Steve - BBC)  $\sigma_w$           & $0.057$ & $0.062$ & $0.060$ \\
		\hline
	\end{tabular}
\end{table}

>>>>

% Abstract of the paper
\begin{abstract}
We present a hierarchical Bayesian model for use in supernova cosmology which takes into account a full range of analysis systematics and selection effects. This advances previous works with hierarchical Bayesian models by including additional sources of systematic uncertainty and improving treatment of Malmquist bias, whilst increasing numerical efficiency. Our model is optimised for efficient fitting and has been validated on a combined simulated dataset of more than $250\,000$ supernovae to rule out the presence of significant model biases in parameter recovery. We demonstrate its effectiveness by fitting simplified statistical simulations for ideal supernova cosmology, and sophisticated SNANA simulations configured to mimic the Dark Energy Survey (DES) 3-year spectroscopic supernova sample. SNANA simulations can recover the input cosmology between $0$ and half the statistical uncertainty from a DES 3-year spectroscopic sample. The method fits cosmology, population and systematics simultaneously while taking into account an increased number of systematics and minimising simulation dependence. Future development will focus on increasing model agnosticism towards intrinsic scatter model.
\end{abstract}
% Bias in SNANA tests across intrinsic dispersion models is found to range from $-0.15\sigma$ to $+0.27\sigma$ in $w$ for datasets of comparable strength to the Dark Energy Survey three-year analysis.

@@

Whilst supernova observations take the form of time-series photometric measurements of brightness in many photometric bands, most analyses do not work from these measurements directly. Instead, most techniques fit an observed redshift and these photometric observations to a supernova model, with the most widely used being that of the empirical SALT2 model \citep{Guy2007, Guy2010}. This model is trained separately before fitting the supernova light curves for cosmology \citep{Betoule2014, Scolnic2017}. The resulting output from the model is, for each supernova, a characterised amplitude $x_0$ (which can be converted into apparent magnitude $m_B = -2.5\log(x_0)$), a stretch term $x_1$ and colour term $c$, along with a covariance matrix describing the uncertainty on these summary statistics, $\mathcal{C}$. As all supernova are not identical, an ensemble of supernovae form a redshift-dependent, observed population of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$, where the hat denotes an observed variable.

@@

Whilst there are a large number of hierarchical models available, none of them have undergone high-statistic tests using realistic simulations to verify each models' respective bias. Additionally, tests that have been done have generally been performed on either  $\Lambda$CDM cosmology or Flat $\Lambda$CDM cosmology and not $w$CDM cosmology. However, quantifying the biases on $w$CDM cosmology with realistic simulations is becoming critically important as precision supernovae cosmology comes into its own and focus shifts from determination of $\Omega_m$ to $w$.

@@

Forefront among these challenges is our ignorance of the underlying Type Ia population dispersion. Ideally, analysis of the underlying population would make use of an ensemble of time-series spectroscopy to characterise the diversity of Type Ia supernovae, however this data is difficult to obtain, and recent efforts to quantify the population draw inference from photometric measurements. We utilise two dispersion models in this work. The first is the \citet[][hereafter denoted {\gten}]{Guy2010} scatter model, which models intrinsic scatter with a 70\% contribution from coherent variation in the spectral energy distribution and 30\% from chromatic variation. The second, denoted the {\celeven} model, is sourced from \citet{Chotard2011} and has variation with 25\% contribution from coherent scatter and 75\% from chromatic variation. As the SALT2 model does not include full treatment of intrinsic dispersion, each scatter model results in different biases in $m_B$, $x_1$ and $c$ when fitting the SALT2 model to light curve observations, and results in increased uncertainty on the summary statistics that is not encapsulated in the reported covariance $\mathcal{C}$. These two scatter models are currently considered sufficient to span the possible range of scatter in the underlying supernova population. We have insufficient information to prefer one model over the other, and thus we have to account for both possible scatter models.

@@

\begin{table}
	\centering
	\caption{ Parameters defined in our model and a summary of their use. The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova. The bottom section shows the observables (not parameters) that are the input data to the model.}
	\label{tab:param_summary}
	\begin{tabular}{ll}
		\hline
		Parameter & Description \\
		\hline
		$\Omega_m$  & Matter density  \\
		$w$  &  Dark energy equation of state  \\
		$\alpha$ & Stretch standardisation   \\
		$\beta$         &  Colour standardisation   \\
		$\delta(0)$ & Scale of the mass-magnitude correction\\
		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
		$\delta\mathcal{Z}_i$ & Systematics scale\\
		$\langle M_B \rangle$ & Mean absolute magnitude \\
		%$s_c$ & Colour bias\\
		\hline
		$\delta S$ & Selection effect deviation \\
		$\langle x_1^i \rangle$ & Mean stretch nodes\\
		$\langle c^i \rangle$ & Mean colour nodes\\
		$\alpha_c$ & Skewness of colour population \\
		$\sigma_{M_B}$ & Population magnitude scatter \\
		$\sigma_{x_1}$ & Population stretch scatter \\
		$\sigma_{c}$ & Population colour scatter \\
		$\kappa_{0}$ & Extra colour dispersion\\
		$\kappa_{1}$ & Redshift-dependence of extra colour dispersion\\
		\hline
		$m_B$ & True flux\\
		$x_1$ & True stretch \\
		$c$ & True colour \\
		$z$ & True redshift \\
		$M_B$ & Derived absolute magnitude \\
		$\mu$ & Derived distance modulus \\
		\hline
		$\hat{m}_B$ & Measured flux\\
		$\hat{x}_1$ & Measured stretch \\
		$\hat{c}$ & Measured colour \\
		$C$ & Covariance on flux, stretch and colour \\
		$\hat{z}$ & Observed redshift \\
		$\hat{m}$ & Determined mass probability \\
		$\hat{p}$ & Determined Ia probability \\
		\hline
	\end{tabular}
\end{table}

@@

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. We note that extra dispersion in magnitude (from coherent scatter) is absorbed completely by the width of the underlying magnitude population (discussed in Section \ref{sec:underlying}) without introducing cosmological bias, which is not true of the colour term, hence the requirement for modelling additional colour dispersion. Tests on incorporating extra dispersion on stretch as well show that stretch is less biased than colour, and cause negligible bias in cosmology.

As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence, increasing with redshift. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature). The $\kappa$ parameters are highly degenerate with the width of the intrinsic colour population $\sigma_c$. We subject them to Cauchy priors centered on zero and with width $0.05$, where $\kappa$ is bounded between $0$ and $0.05$. We pick this maximum value to allow extra dispersion without completely subsuming the intrinsic population widths due to the severe degeneracy, where this maximum value easily encapsulates the determined dispersion according to the results of \citet{Kessler2013}. As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.

Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ (rather than just dispersion on $c$) was also tested, by modelling the dispersion as a multivariate Gaussian, but it showed negligible improvement in recovering unbiased cosmology over just colour dispersion, and was far more computationally inefficient. We note here that we do model dispersion in magnitude, but this is done at the level of underlying populations, not observed populations. This extra dispersion is modelled with redshift independence.

@@

We tested two models with varying $\beta$ against simulated supernova sets; $\beta(c) = \beta_0 + \beta_1 c$ and $\beta(z) = \beta_0 + \beta_1 z$. See Section \ref{sec:simdes} for details on simulation generation. We found for both models that non-zero values for $\beta_1$ are preferred (even with constant $\beta$ used in simulation) due to severe degeneracy with selection effects. This degeneracy resulted in a significant bias in recovered cosmology. Due to the recovery of non-zero $\beta_1$, we continue to adopt the constant $\alpha$ and $\beta$ found in traditional analyses. As such, our calculation of distance modulus $\mu$ mirrors that found in Equation \eqref{eq:mu0}.

@@

Early analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}.

@@

In addition to the improvements in the scatter models over the simple data, we also include peculiar velocities for the LowZ sample, and our full treatment of systematics. Our simulated populations are sourced from \citet[][hereafter {\sk}]{Scolnic2016} and shown in Table \ref{tab:dist}. Initial tests were also done with a second, Gaussian population with colour and stretch populations centered on zero and with respective width $0.1$ and $1$, however cosmological parameters were not impacted by choice of the underlying population and we continue using only the {\sk} population for computational efficiency. The selection effects were quantified by comparing all the generated supernovae to those that pass our simulated cuts, as shown in Figure~\ref{fig:sf_snana}. It is from this simulation that our analytic determination of the selection functions for the LowZ and DES survey are based. We run two simulations to determine the efficiency using the {\gten} and {\celeven} scatter models and find no difference in the Malmquist bias between the two models.

@@

Points of maximum posterior for 100 data realisations are shown in Figure \ref{fig:bulk_posterior}. The parameter bounds and biases for $w$ are listed in Table \ref{tab:bulk_summary},  and secondary parameters are shown in Table \ref{tab:standardisation}. Our recovered results for $w$ are all within $0.5\sigma$ of the true values of $w=-1$. Table \ref{tab:standardisation} shows a clear difference in both $\beta$ and $\sigma_{m_B}$ across the {\gten} and {\celeven} simulations. As expected, the {\celeven} simulations recover a far smaller intrinsic magnitude scatter, giving results of approximately $0.025$ when compared to the result of $0.070$ for the {\gten} simulations. The extra smearing of the {\celeven} model does not result in a biased $\beta$ value, with recovery of $\beta\approx3.76$ close to the input truth value of $3.8$, however the $\beta$ recovery for the {\gten} simulations is biased high, finding $\beta\approx3.44$ with an input of $3.1$. Interestingly, $w$-bias is only found for the {\celeven} simulations.

We investigate the cosmological bias and find its source to be a bias in the observed summary statistics, in addition to incorrect reported uncertainty on the summary statistics. To confirm this, we run two tests. The first of which, we replace the fully simulated observed $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ with random numbers drawn from a true Gaussian centered on the simulated SALT2 $m_B$, $x_1$ and $c$ values with covariance as reported by initial light curve fits. With this test, both the {\gten} and {\celeven} fits recover $w=-1.00$ exactly. Our second test, to allow measurement biases not sourced from intrinsic scatter through, we test a set of 100 simulations generated using only coherent magnitude scatter, and also find $w=-1.00$, showing that the source of the biases in summary statistics is the underlying intrinsic scatter model. From this, the main challenge of improving our methodology is to handle the fact that observational uncertainty reported from fitting the SALT2 model to light curves is incorrect, non-Gaussian and biased. Our current model and techniques can quantify the effect of different scatter models on biasing the observed summary statistics, but being unable to constrain the `correct' (simulated) scatter model in our model fit means we cannot fully correct for the bias introduced an unknown scatter model.

Unfortunately, adding extra fit parameters to allow for shifting observables washes out our ability to constrain cosmology, and applying a specific bias correction requires running a fiducial simulation (assuming cosmology, population and scatter model), which presents difficulties when trying to account for correlations with population and scatter model. This is compounded by the fact that bias corrections do not in general improve fits (increase the log posterior), and so are difficult to fit parametrically. Works such as \citet{Kessler2017} show that bias corrections can be applied to supernovae datasets that can robustly handle multiple intrinsic scatter models, and future work will center on uniting these methodologies - incorporating better bias corrections that separate intrinsic scatter bias and non-Gaussian summary statistic bias from Malmquist bias, without having to precompute standardisation parameters and populations.

@@

Table \ref{tab:parameter_correlations} lists the fit correlations between our model fit parameters (excluding the LowZ band systematics, and Malmquist bias uncertainty parameters which had negligible correlation), showing (in order) cosmological parameters, standardisation parameters, population width and skewness parameters, intrinsic dispersion parameters, mass-step parameters, population mean parameters, SALT2 model systematics, dust systematic, global HST calibration systematic, peculiar velocity systematic, global redshift systematic and DES band magnitude and wavelength systematics. Figure \ref{fig:bulk_cor} show the full correlations between all non-systematic model parameters. Other interesting correlations are shown and discussed in Figure \ref{fig:bulk_cor}. The band systematics for DES filters $g$, $r$ and $i$ also show significant correlation with $w$, highlighting the importance of minimising instrumental uncertainty.

@@

\begin{table*}
	\centering
	\caption{The error budget on $w$, as determined from analysing uncertainty on simulation data whilst progressively enabling model features. We start from the top of the table, only varying cosmological parameters $\Omega_m$ and $w$, and then progressively unlock parameters and let them fit as we progress down the table. The cumulative uncertainty shows the total uncertainty on $w$ on the fit for all, where the $\sigma_w$ term is derived by taking the quadrature difference in cumulative uncertainty as we progress.}
	\label{tab:uncert}
	\begin{tabular}{l|l|rr}
		Feature &  Parameters & $\sigma_w$ & Cumulative  \\
		\hline
		Cosmology                      & $\Omega_m$, $w$ & 0.051  &   0.051       \\
		Standardisation                & $\alpha$, $\beta$, $\langle M_B \rangle$, $\delta(0)$, $\delta(\infty)/\delta(0)$ & 0.046  &   0.068       \\
		Intrinsic scatter              & $\kappa_0$, $\kappa_1$ & 0.020  &   0.071       \\
		Redshift-independent populations             & $\sigma_{M_B}$, $\sigma_c$, $\sigma_{x_1}$, $\alpha_c$ & 0.022  &   0.074       \\
		Redshift-dependent populations & $\langle c_i \rangle$, $\langle x_{1,i} \rangle$ & 0.030  &   0.080   \\
		Systematics                    & $\delta \mathcal{Z}_i$, $\delta S$ & 0.054  &   0.096      \\ 
		\hline
	\end{tabular}
\end{table*}

@@

As shown in \citet{Brout18SYS}, BBC recovers cosmological parameters without bias so long as the intrinsic scatter model is known. As we do not know the correct intrinsic scatter model, the BBC method averages the results when using bias corrections from {\gten} and from {\celeven}. As such, we expect the BBC method to be biased in one direction for {\gten} simulations and the other direction for {\celeven} simulations. These results are consistent with those displayed in Table \ref{tab:bulk_summary_bbc}, as both BBC and BHM are sensitive to the intrinsic scatter model, finding differences of $\sim0.06$ in $w$ when varying the scatter model. The BBC method finds $w$ biased low for G10 and $w$ biased high for C11 (by about $\pm0.03$), so taking the average result only results in a small bias of $-0.01$ in $w$. Our method shows a small improvement in the insensitivity to the intrinsic scatter model, finding no bias for {\gten} but a $w$ biased high for {\celeven}. This decrease in error is not statistically significant as we have statistical uncertainty of $\sim0.01$ for $100$ simulation realisations. The average bias over the two scatter models is $+0.028$, representing a larger bias than the BBC method.

@@

\begin{table}
	\centering
	\caption{Characterising the bias on $w$ using the 100 simulations for the {\gten} scatter model and 100 simulations for {\celeven} scatter model. The mean $w$ value for our method and BBC are presented, along with the mean when averaging the difference between our method and BBC for each individual simulation. Finally, we also characterise the scatter between the methods in the final row. Averages are computed giving each simulation sample the same weight. In the model, $\Delta$ represents Steve - BBC. The final row shows the scatter between Steve and BBC for the different simulations.}
	\label{tab:bulk_summary_bbc}
	\begin{tabular}{l|ccc}
		\hline
		Model & {\gten} & {\celeven} & ({\gten} + {\celeven}) \\
		\hline
		Steve $\langle w \rangle$           & $-0.998\pm0.007$ & $-0.945\pm0.007$ & $-0.972\pm0.006$\\
		BBC $\langle w \rangle$             & $-1.044\pm0.006$ & $-0.978\pm0.007$ & $-1.010\pm0.005$ \\
		$\Delta$  $\langle w \rangle$  & $+0.044\pm0.006$ & $+0.033\pm0.006$ & $+0.038\pm0.004$ \\
		\hline
		$\Delta$  $\sigma_w$           & $0.057\pm0.004$ & $0.062\pm0.004$ & $0.060\pm0.003$ \\
		\hline
	\end{tabular}
\end{table}
